Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-07 doc/report/msc-report.tex
wrapped the lines of that text I wrote earlier today.
added a 500 word section on the implementation of dimensionality reduction, etc

>>>>

- Something about cardinality -

When reasoning about dimensions, the issue of cardinality will come up sooner
or later.

The cardinality of a dimension is the number of distinct values of the
dimension. For example, the dimension of 'logged in' has a cardinality of two
since 'logged in' is a boolean variable; the user is either logged in or not.
The 'country' dimension has a much higher cardinality: there are some 300
countries in the world. Certainly, there could be dimension whose cardinality
is infinite: the 'path' dimension in the web log example, for example, could
in theory have a limitless number of values.
The number of possible combinations of dimensions of your data -- the number
of points in n-space -- is directly related to the cardinality of each
dimension. It is likely that your data is of such a nature that it is
physically impossible to store information about all dimension combinations.
This is why we ask you, the user, to define what projections you are
interested in: storing aggregates for the slices of the cube that interest you
is both a time and space save, and in many cases the only feasible solution.

-- Limits on cardinality --

[What are the limits of our system? How high cardinality do we support? (This
is closely tied to hbase and the qualifier-limit.) If Martin is going to have
'normalized_path' as one of his dimensions, we might run into trouble.]

-- The implementation of all that --

Since we wish to store projections rather than all combinations of all
dimensions, we need to mangle the data slightly before sending it down the
pipe to the underlying data store.

Dumbo expects a key-value pair from the mapper. Hadoop will collect all keys
and give the reducer a key and an iterator of values. If we are to store
projections, we need to encode both the desired dimensions and the unit in
question in the key and emit the value of the unit in the 'value' field. This
way we can use a simple SumReducer to sum the values for each unit at any
point in n-space and be done.

Remember, the usermapper gives us a point in n-space and a list of
measurements. We need to break out the measurements and sum over each such
measurement. Zohmg solves this by wrapping the usermapper in a cleverly named
devicie, namely a mapperwrapper.

% TODO: this para sounds as if it was written by a drunken monkey. The
mapperwrapper takes each tuple emitted by the usermapper and performs a
dimensionality reduction, ending up with the projections requested by the
user. For every projection and unit, the dimensions remaining after the
reduction are appended to the string representation of the unit and emitted as
the key. The value is simply the value of the unit. The Hadoop framework will
do the heavy work of collecting all keys and summing the values.

An alternative approach would have been to perform the dimensionality
reduction at a later stage, namely in the reducer of a multi-pass MapReduce
job.

The reducer, then, is very simple: it's a simple sum of the values. The output
of the reducer will not be stored on disk as is usually the case, but will
instead be interpreted by a custom OutputReader that persists the data to
HBase.

The reducer needs to communicate to the OutputReader under what rowkey,
column-family and qualifier to store the computed sum. It does this by
emitting the row key as the key and a json-serialized map of the form
{"column-family:qualifier" : value}.

The OutputReader interprets the json and persists the data. Et voila!
